\section{Online Coloring Co-Interval Graphs}
This section studies the online problem of coloring co-interval graphs as described by Zarrabi-Zadeh \cite{zarrabi}.
This problem is set up as follows:

\begin{description}
	\item[Input] A set of intervals on the real line in arbitrary order.
	\item[Output] A color matched to each interval, such that $color(I_k) \neq color(I_j)$ when intervals $I_k$ and $I_j$ are such that $j < k$ and intervals $I_j$ and $I_k$ don't intersect.
\end{description}

Since this is an online problem, the algorithm must assign colors as it goes, without knowing the complete set of intervals. That is, the algorithm will only receive information about $I_{j+1}$ after it has assigned colors to all intervals $I_k$ where $j \geq k$.

First, we present the First-Fit approximation algorithm, which solves this problem with a competitive ratio of at most 2. Then, using Yao's principle, we can prove that there exist no randomized algorithm that achieves a competitive ratio that's better than $\frac{3}{2}$.

\subsection{Upper Bound}
To solve the problem of assigning intervals different colors, the set of intervals is first constructed as a graph. That is, the input to our problem becomes:

\begin{center}
	\begin{tabular}{rl}
		$V$ &:= Set of intervals where $v_i$ represents the $i$\textsuperscript{th} interval \\
		$E$ &:= Set of edges where $(u,v) \in E$ iff interval $u$ and interval $v$ are disjoint
	\end{tabular}
\end{center}

Note that any two vertices that aren't connected by an edge can be colored in the same color, since those vertices represent overlapping intervals. Thus, in such a graph, our problem is reduced to finding the minimum number of colors such that no connected vertices are the same color.

The First-Fit algorithm, as follows, is the simplest solution to this problem.

\begin{algorithm}
\caption{First-Fit algorithm}
$L \leftarrow$ ordered list of colors\;
\ForEach{new vertex v} {
  assign $v$ the first color in $L$ that doesn't cause a collision\;
}
\end{algorithm}

Essentially, the algorithm maintains a list of colors that can be used. These colors are sorted in an arbitrary order. The first vertex is simply assigned the smallest color. For any future vertex $v$, the algorithm scans its list of colors. A \emph{collision} between vertices $u$ and $v$ at color $k$ occurs if $u$ is colored $k$ and $v$ is connected to $u$ by an edge in $E$. If $k = L[i]$ and a collision occurs, then we increment $i$. We assign $v$ the color at $L[j]$, where $j$ doesn't cause any collisions, and we try to minimize $j$.

It has been shown by Gyarfas and Lehel \cite{gyarfas} that First-Fit uses at most $2\omega+1$ colors on any co-chordal graph, which is a class that includes co-interval graphs. It has also been shown by Kierstead and Qin \cite{kierstead} that no online deterministic algorithm can beat the 2-competitive ratio on co-interval graphs.

In the next section, we look at using randomization in co-interval graphs and lower bound the best competitive ratio that can be obtained.
\subsection{Lower Bound}
Yao's minimax principle proves that the expected competitive ratio of an optimal deterministic algorithm on the worst-case distribution of inputs is a lower bound on the expected competitive ratio of all randomized algorithms. That is, if there exists a set of inputs such that all optimal deterministic algorithms are at least $\alpha$-competitive, then $\alpha$ is the best ratio that can be obtained by using randomization.

To define such a set of inputs, essentially, we need to find a set of intervals that satisfy the condition that any deterministic algorithm is at most $\frac{3}{2}$-competitive. The idea behind constructing such a set of intervals is that any deterministic algorithm will make the same decision on any specific interval given the same information, regardless of what the future intervals are.

\subsubsection{Problem Construction}
For $1\le i\le k$ for a given value of $k$, the intervals defined by Zarrabi-Zadeh \cite{zarrabi} are: 
\begin{align*}
	a_i &= [3i-3, 3i-2] \\
	b_i &= [3i+1, 3i+2] \\
	c_i &= [3i+2, \infty]
\end{align*}
We can then define a block as:
$B_i = \begin{cases} {b_1, c_1} &\mbox{if } i = 1 \\
{a_i, b_i, c_i} & \mbox{if } i \in [2,k) \\
 {a_k, b_k} &\mbox{if } i = k \end{cases}$ 

From this set of blocks, we can choose a set of $k$ input sequences, simply by defining our input sequence $X_i$ as the concatenation of blocks $B_1$, $B_2$... $B_i$, in order. Let $I$ be our set of all possible inputs:

\[I = \{X_i \text{ where } i \in [1,k]\}\]

It can be shown that the minimum number of colors required to color $X_i$ without any edges sharing a color is $i$.

To use Yao's principle on this input distribution, we need to show that the expected competitive ratio will be at least $\frac{3}{2}$ if we set up a probability distribution over all possible $X_i$. The optimal deterministic algorithm that we will examine per each $X_i$ input is First-Fit; as previously shown, this algorithm uses $2\omega-1$ colors on each $X_i$. In this case, the chromatic number $\omega=i$, so, the expected number of colors for a given $X_i$ is exactly $2i-1$.

For a given $X_i$, a new color is chosen (assuming First-Fit) on each $a_m$ and each $b_m$, since the blocks are given in order. A probability distribution $P$ can be defined over all $X_i$, where $p_i$ is the probability that $X_i$ is chosen as input, as follows:

\[p_i = \begin{cases} \frac{2^k}{2^k-1}\cdot\frac{i}{2^i} &\mbox{if } i < k \\
\frac{k}{2^k-1} & \mbox{if } i \equiv k 
 \end{cases}\]

The expected competitive ratio of First-Fit on this distribution then becomes $\sum_{i=1}^k p_i\frac{2i-1}{i}$, since First-Fit uses $2i-1$ colors, whereas the optimal number of colors that can be used is exactly $i$.

This sum becomes:
\begin{equation*}
\rho = \frac{3}{2} - \frac{1}{2(2^k-1)}
\end{equation*}

where $\rho$ is the best competitive ratio that can be obtained by using First-Fit on this probability distribution. 

If no deterministic algorithm can obtain a better competitive ratio on this set of inputs, then this is the best that a randomized algorithm could achieve.
\subsubsection{Deterministic Algorithm Bounds}
\begin{theorem}
	No deterministic algorithm can obtain a better competitive ratio on the distribution of inputs $I$.
\end{theorem}
\begin{proof}
To show this, we choose an arbitrary deterministic algorithm $A$. We can then define $d_i$ as the decision that $A$ makes on receiving interval $c_i$, where $d_i = 1$ iff a new color is chosen for $c_i$.  In an optimal solution, all $c_i$ can be colored the same as $b_i$, since all $b_i$ don't intersect any preceding $a_j$ or $b_j$, but intersect all previous $c_j$.  

Given a sequence of decisions, $\langle d_1, d_2...d_{k-1} \rangle$, then we can define a deterministic algorithm FF as follows:

If we receive an interval $c_j$, we check what $d_j$ is and assign a new color iff $d_j = 1$. Else, we assign it the same color as $b_j$. 

If we receive an interval of type $a$ or $b$, we behave as the online First-Fit algorithm normally does.

\begin{lemma} Among all deterministic online algorithms, FF behaves optimally and chooses the minimum number of colors for any input $I$. \end{lemma}
\begin{proof}
This can be shown by induction over the minimum number of colors on an input of size $m$. 
%%Then, assuming that the minimum number of colors are used in an input of size $m$, then we know that all deterministic algorithms will behave the same way for a specific prefix of size $m$. We can then show that the minimum number of colors will be used on an input of size $m+1$ and the statement holds.
\end{proof}

\begin{lemma} The competitive ratio of FF is $\rho$. \end{lemma}
\begin{proof}
This proof follows by induction:

Define $P(m)$ to be true if $\rho$ is an upper bound for the competitive ratio of FF in the case where there are $m$ 1's in our sequence of decisions $D$.  $P(0)$ holds trivially -- in a sequence of all 0's, FF behaves exactly the same as First-Fit and our competitive ratio is still \textasciitilde2.

Using strong induction, we assume that $P(k)$ is true for all $k \in [0,m-1]$.  

%Then, $i$ is defined as the index of the last bit which is 1 in $D_m$, an arbitrary sequence with $m$ 1's.  If we flip $D_m[i]$, we get a sequence $D_{m-1}$ with a competitive ratio at most $\rho$ (by induction).

%Thus, for the first $i$ bits, the competitive ratio of $FF_m$ is the same as that of $FF_{m-1}$ for all sequences smaller than $X_i$. 

%On sequence $X_i$, $FF_m$ uses one more color than $FF_{m-1}$, since it uses a new color for $c_i$.

%On sequences larger than $X_i$, $FF_m$ uses one less color than $FF_{m-1}$. This is because it opens a new color for $c_i$. However, $FF_m$ colors $a_{i+1}$ like $b_i$ and $b_{i+1}$ like $c_i$. Meanwhile $FF_{m-1}$ opens new colors for $a_{i+1}$ and $b_{i+1}$ since that's what First-Fit does. 

By looking at the bits of $D$ where $FF$ with $m$ 1's would differ from $FF$ with $m-1$ 1's, we can derive the formula that:

\[\rho_m = \rho_{m-1} + \rho_i\left(\frac{1}{i}\right) - \sum_{j=i+1}^k \rho_i\left(\frac{1}{i}\right)\]

Substituting $\rho_i = \frac{2^k}{2^k-1}\cdot\frac{i}{2^i}$ gives us that the last term sums to exactly $\rho_i(\frac{1}{i})$.  This gives us that $\rho_m = \rho_{m-1}$, and the proof follows by induction.
\end{proof}

In turn, this implies that the expected competitive ratio of any deterministic algorithm is at most $\rho$, which shows that no deterministic algorithm has a better competitive ratio on $I$. 
\end{proof}

By choosing $k$ arbitrarily large, $\rho$ tends to $\frac{3}{2}$. As we have now bounded the behavior of any deterministic algorithm on the distribution of inputs $I$, Yao's principle thus gives us a lower bound of $\frac{3}{2}$ as the competitive ratio on any randomized algorithm.
%\subsection{Remarks}
%This is an especially interesting application of Yao's principle because in unit co-interval graphs, randomization can prove a definite improvement in the competitive ratio. That is, it is still unknown whether there exists a deterministic algorithm that is better than 2-competitive in the case of unit co-interval graphs. However, using randomization, we can obtain a $3/2$ competitive ratio.

%Similarly, it is unknown whether we can actually extend the $3/2$ competitive ratio to general co-interval graphs. However, Yao's principle shows us that if this ratio is obtained, it must be tight, without ever analyzing the effects of randomization on the problem itself and just looking at how different deterministic algorithms behave.
